Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2017-02-07 manuscript/editor_response_tree_physiology.Rmd
editors response comments made

\
*Response*: We agree with the editor's comments that the effects of soil volume have not been properly accounted for in several places in the manuscript. Particularly in the discussion we have clarified the nature of the effect type, where soil volume effects have often been over-emphasized. We agree that the stats are strongly influenced by the Free seedling, but this is also representative of the observed effects of root restirction on *A*~sat~, TNC and leaf N across all containers throughout all measurement campaigns. As the impacts of root restriction and poor quality soil impacted were evident across all treatments 
We feel that addressing individual effects of specific volume treatments, or combining soil volume treament effects, would overly complicate the paper and would not improve our intrepretation of the results.

@@

\
*Response*: The editor makes several good points here. We have re-written the equation to better descibe the way daily total leaf net C gain (*P*~day~) was calculated. We have also removed the statement "Leaf respiration is included in the calculation of *P*." and added a accurate description of how the area based respiration component was calculated at each 15 minute interval to then model leaf level net photosynthesis rates.

@@

\
*Response*: We agree with each of the editors points here.  With a comment #6 we have now clearly defined the "total net leaf C gain term".  As the edigor points out there are several places where leaf *A*~n~ is discussed inappropriately, especially in the discussion.  Throughout the manuscript we have revised use of terms for C gain to accurately reflect either measured photosynthesis rates (*A*~sat~), leaf level rates of net photosynthesis (*A*~n~) or modelled cumulative total net leaf C gain (*P*~day~).

@@

\
*Response*: 
\
- *Comment: 18. L413: what is An here? Leaf area or whole plant based?*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat (measured values).  
\
- *Comment: 19. L430-431. Correlation between leaf N and An (what is it?) has not been shown in the result sections.*

\
*Response*: We have corrected *A*~n~ to *A*~mass~, which is reported in the results section.
\
- *Comment: 20. L434: the connexion between root N and N uptake is not trivial.*

\
*Response*: We agree with the editor. We have changd this sentence to give a quantative description of root N content differences between roots of Free seedlings and seedlings in containers (+15 %). 
\
- *Comment: 21. L441: You did not measure available soil N pool but total soil N, which is different. It would have been interesting to know if microbial N mineralization is affected by the containers or not. The key might not be at the plant level but at the microbe level. And what about other nutrients such as available P?*

\
*Response*: We agree with the editor's comments. There is convincing evidence that nutrient uptake was negatively affected in this experiment but the mechanism is yet unidentified. Understanding the processes by which belowground sink manipulation manifests belowground will be important for future expereiments using local soils. We hope that this dataset and experimental design can aid in future experiments to further investigate the mecanisms underpinning belowground sink limitation.

@@

\
*Response*: We have corrected this statment in the discussion to more accurately reflect the observed results.

@@

\
*Response*: We have added both measured harvested leaf area and *A*~sat~ into Table 1 and 3, respectively.  Both *A*~sat~ and leaf area were significantly reduced in seedings lin containers compared to the free seedlings.

>>>>

\
*Response*: We agree with the editor's comments that the effects of soil volume have not been properly accounted for in several places in the manuscript. Particularly in the discussion we have clarified the nature of the effect type, where soil volume effects have often been over-emphasized. We agree that the stats are strongly influenced by the Free seedling, but this is also representative of the observed effects of root restriction on *A*~sat~, TNC and leaf N across all containers throughout all measurement campaigns. As the impacts of root restriction and poor quality soil impacted were evident across all treatments 
We feel that addressing individual effects of specific volume treatments, or combining soil volume treatment effects, would overly complicate the paper and would not improve our interpretation of the results.

@@

\
*Response*: The editor makes several good points here. We have re-written the equation to better describe the way daily total leaf net C gain (*P*~day~) was calculated. We have also removed the statement "Leaf respiration is included in the calculation of *P*." and added a accurate description of how the area based respiration component was calculated at each 15 minute interval to then model leaf level net photosynthesis rates.

@@

\
*Response*: We agree with each of the editors points here.  With a comment #6 we have now clearly defined the "total net leaf C gain term".  As the editor points out there are several places where leaf *A*~n~ is discussed inappropriately, especially in the discussion.  Throughout the manuscript we have revised use of terms for C gain to accurately reflect either measured photosynthesis rates (*A*~sat~), leaf level rates of net photosynthesis (*A*~n~) or modelled cumulative total net leaf C gain (*P*~day~).

@@

\
*Response*: We have revised this statement to make the point that reductions in measured *A*~sat~ and seedling allometry occurred simultaneosly during early measurement campaigns, and these reductions where sustained throughout the experiement.

\
- *Comment: 18. L413: what is An here? Leaf area or whole plant based?*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat (measured values).

\
- *Comment: 19. L430-431. Correlation between leaf N and An (what is it?) has not been shown in the result sections.*

\
*Response*: We have corrected *A*~n~ to *A*~mass~, which is reported in the results section.

\
- *Comment: 20. L434: the connexion between root N and N uptake is not trivial.*

\
*Response*: We agree with the editor. We have changed this sentence to give a quantitative description of root N content differences between roots of Free seedlings and seedlings in containers (+15 %). 
\
- *Comment: 21. L441: You did not measure available soil N pool but total soil N, which is different. It would have been interesting to know if microbial N mineralization is affected by the containers or not. The key might not be at the plant level but at the microbe level. And what about other nutrients such as available P?*

\
*Response*: We agree with the editor's comments. There is convincing evidence that nutrient uptake was negatively affected in this experiment but the mechanism is yet unidentified. Understanding the processes by which belowground sink manipulation manifests belowground will be important for future experiments using local soils. We hope that this data-set and experimental design can aid in future experiments to further investigate the mechanisms underpinning belowground sink limitation.

@@

\
*Response*: We have corrected this statement in the discussion to more accurately reflect the observed results.

@@

\
*Response*: We have added both measured harvested leaf area and *A*~sat~ into Table 1 and 3, respectively.  Both *A*~sat~ and leaf area were significantly reduced in seedlings in containers compared to the free seedlings. Leaf *A*~n~ was also replaced with "photosynthetic parameters".

### 2017-02-07 manuscript/manuscript_pve_editor.Rmd
editors response comments made

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through net photosynthesis must be balanced by C allocation to growth and storage and C losses such as respiration, root exudation, volatile organic compounds and tissue turnover. Plant growth is thus predicted as the outcome of the processes of photosynthesis, C allocation and C loss. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, C allocation and C loss. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

\
Leaf dark respiration rates (*R*) were measured on each seedling on the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions (11.5-13.4 &deg;C). *R* was logged 5 times for each leaf after CO~2~ and water vapour flux values stabilized in the conifer chamber. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment (Drake *et al.* unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

@@

\
(1)
$$G(t) = \sigma LP_{day}$$

where $\sigma$ is a self-shading parameter, *L* is modelled total plant leaf area (m^2^), and  *P*~day~ is the net daily C assimilation per unit leaf area (g C m^-2).

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *g*~1~ and *R*~dark~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above). The *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. *R*~dark~ was was estimated using Eq. (2) from @atkin2005response, which was parameterized with a mean reference empircal measurement for each treatment, a Q~10~ value of 1.86 from a seperate experiment using the same tube stock seedlings, a reference air temperature and air temperatures at each time step.

\
(2)
$$R~dark~ = R_{ref}Q_{10}~^{(T-T_{ref})/10}$$
Methods of the coupled leaf gas exchange model are also described in Duursma *et al.* [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A*~n~ rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *P*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *P* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *P*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

\
We used modelled C gain  to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A*~n~ were sufficient to accurately predict overall seedling biomass production after 120 days. Cumulative net leaf C gain for each treatment was equal to the sum of each value of *P* over 120 days and final seedling C was assumed to equal half of the final dry mass for observed seedlings. We then compared seedling C use efficiency (CUE) across treatments, as the proportion of modelled cumulative net leaf C gain allocated to observed seedling biomass.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis (*A*~mass~) was correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~mass~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~mass~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~mass~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling growth. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~sat~ was also reduced, but this reduction was insufficient to fully explain the reduction in growth, implying that components such as non-leaf respiratory losses were also likely affected.

@@

\
Decreases in *A*~sat~ occurred simultaneously with reductions of growth of seedlings in containers. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~sat~ in small containers including increased soil temperatures, reduced water availability or nutrient content, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine the mechanism by which the induced belowground sink limitation triggered photosynthetic down regulation. In this study, the impacts of increasing soil temperatures with decreasing container sizes was controlled by burying of all containers flush to the soil surface.

@@

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~mass~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza *et al.* [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of Free seedlings in our study, this suggests the response of *A*~n~ to sink inhibition was regulated by this accumulation, as hypothesized. Overall, it is likely that both nutrient limitation and reduced sink strength played a role in observed photosynthetic down regulation, but future studies are needed to identify the specific mechanisms which prompted these feedbacks.

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf *A*~n~ , via soil volume effects.  Although reductions in both *A*~max~ and biomass were correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As estimated whole-plant C use efficiency was decreased in containerized seedlings, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].

>>>>

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through net photosynthesis must be balanced by C allocation to growth and storage and C losses including respiration, root exudation, volatile organic compounds and tissue turnover. Plant growth is thus predicted as the outcome of the processes of photosynthesis, C allocation and C loss. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, C allocation and C loss. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

\
Leaf dark respiration rates (*R*) were measured on each seedling on the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions (11.5-13.4 &deg;C). *R* was logged 5 times for each leaf after CO~2~ and water vapour flux values stabilized in the conifer chamber. Reported values of *R* are standardized rates at 25 &deg;C using a *Q*~10~ value (1.86) developed for these seedlings in a separate experiment. Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

@@

\
(1)
$$G(t) =  L \sigma P_{day}$$

where *L* is modelled total plant leaf area (m^2^), $\sigma$ is a self-shading parameter and  *P*~day~ is the net daily C assimilation per unit leaf area (g C m^-2^).

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *g*~1~ and *R*~dark~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above). The *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. *R*~dark~ was was estimated using Eq. (2) from @atkin2005response, which was parameterized with a mean reference empircal measurement for each treatment, a Q~10~ value of 1.86, a reference air temperature for each treatment and air temperatures at each time step.

\
(2)
$$R~dark~ = R_{ref}Q_{10}~^{(T-T_{ref})/10}$$
Methods of the coupled leaf gas exchange model are also described in Duursma *et al.* [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, net leaf photosynthesis rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. *P*~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of *P* for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of *P*~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

@@

\
We used modelled C gain  to test the hypothesis that the effects of belowground sink limitation on rates of leaf photosynthesis were sufficient to accurately predict overall seedling biomass production after 120 days. Cumulative net leaf C gain for each treatment was equal to the sum of each value of *P* over 120 days and final seedling C was assumed to equal half of the final dry mass for observed seedlings. We then compared seedling C use efficiency (CUE) across treatments, as the proportion of modelled cumulative net leaf C gain allocated to observed seedling biomass.

@@

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis (*A*~mass~) was correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~mass~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~mass~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~mass~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b).

@@

#Discussion
This study utilized a novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling growth. Our goal was to identify how sink limitation of growth modified the C balance of seedlings. We found a strong reduction of growth in containerized plants. Leaf *A*~sat~ was also reduced, but reductions in net leaf photosynthesis rates were insufficient to fully explain the reduction in growth, implying that components such as non-leaf respiratory losses were also likely affected.

@@

\
Decreases in *A*~sat~ occurred simultaneously with reductions of height, diameter and leaf area of seedlings in containers throughout the duration of the experiment. This observation initially suggests that sink limitation may be driving the down regulation of photosynthesis, and thus growth. However, there are several possible mechanisms that can explain reduced *A*~sat~ in small containers including increased soil temperatures, reduced water availability or nutrient content, or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine the mechanism by which the induced belowground sink limitation triggered photosynthetic down regulation. In this study, the impacts of increasing soil temperatures with decreasing container sizes was controlled by burying of all containers flush to the soil surface.

@@

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A*~n~ from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process [@nebauer2011photosynthesis]. Here, declines in *A*~mass~ were correlated with higher starch content throughout the experiment. These results agree with a study on a deciduous conifer by Equiza *et al.* [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of Free seedlings in our study, this suggests the response of *A*~sat~ to sink inhibition was regulated by this accumulation, as hypothesized. Overall, it is likely that both nutrient limitation and reduced sink strength played a role in observed photosynthetic down regulation, but future studies are needed to identify the specific mechanisms which prompted these feedbacks.

@@

##Do reductions in photosynthesis explain reductions in seedling growth?
We use a simple approach to estimate cumulative seedling C gain with measured reductions in leaf photosynthetic parameters, via soil volume effects.  Although reductions in both *A*~max~ and biomass were correlated among treatments, as hypothesized by Poorter *et al.* [-@poorter2012pot], we provide evidence that the negative effects of sink limitation on *A*~n~ do not fully explain reduced seedling growth. As estimated whole-plant C use efficiency was decreased in containerized seedlings, this suggests that other components of the C balance were affected. These results are especially noteworthy for process-based growth models that parameterize CUE, tissue respiration and C allocation as fixed processes. For example, classical approaches in tree growth and production modelling are often driven by inputs of C assimilation and processes such as respiration are considered proportional to biomass [@le2001carbon] or growth rate [@tjoelker1999acclimation]. It is possible these results represent an initial overestimation of *A*~n~, however, the robust empirical based methods used to generate photosynthetic parameters (*J*~max~, *V*~cmax~, *R*~dark~ and *g*~1~) make this unlikely. Instead, our results indicate a need to evaluate how oversimplified representations of processes other than *A*~n~ affect models which distinguish the fate of assimilate C within a plant. Doing so will provide valuable input to future models as assimilate allocation is a key component in C-balance driven plant growth models [@lacointe2000carbon].
